\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-23 draft.tex
Section 2: preliminaries, input language

\begin{document}

@@

\hide{More concretely, for the verification of a MapReduce program, we propose
to do it in two phases. First, ensuring if all reducers are commutative. If some
reducers are non-commutative, modifying them to commutative ones. Usually, the
modification is not a difficult task~\cite{xiao14mr}, and can be done without
affecting their functionality and performance. For example, assume that the task
of a reducer is to find the name of the person with highest score. Such a
reducer is non-commutative when the input include two people with the same
highest score. This reducer can be made commutative, by also comparing the ID
number of people with the same score. Once all reducers are made commutative, we
reduce the verification problem to sequential verification by fixing a
scheduler. In this two-phase approach, the key enabling technique is an
efficient procedure for checking reducer commutativity.}

@@

\hide{
\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}}
\begin{wrapfigure}{r}{3cm}
		\vspace{-0.8cm}
	\hide{
		\begin{minipage}{0.32\textwidth}
			\begin{algorithm}[H]
				$m := \cur$; \;
				\Loop{}{
					$t:=\cur$;\;
					\uIf{ $t > m$}{
						$m := t$
					}
				};
				$out(m)$\;
			\end{algorithm}
			\caption*{(a) max}
		\end{minipage}
		\hspace{-0.5cm}
	}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
	\end{minipage}
	\hide{
		\begin{minipage}{0.4\textwidth}
			\begin{algorithm}[H]
				$a:=*;s := 0; c:=0$;\;
				\Loop{}{
					$s := s+(\cur-a)^2$; \;
					$c := c+1$
				};
				$o := s/c$;\;
				$out(o)$
			\end{algorithm}
			\caption*{(c) standard deviation}
		\end{minipage}
	}
	\caption{Average}
	\vspace{-0.8cm}
	\label{fig:reducer_example}
\end{wrapfigure}
For presentation purpose, we describe a highly simplified language to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.
Let $\Var$ be the set of variables. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. An example is in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the final result. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average and the other reducer performs the rest of the computation.
The second reducer replaces the average with a non-deterministic value and thus ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output for all possible ``average'' value. However, this reducer is still commutative and its commutativity verification can still be done.

\section{Overview of the J-ReCoVer tool}
\label{sec:overview}
%Hadoop, Java;
The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework.
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either through our web-interface or binary application installed on their own machine.

@@

\section{The Preprocessor and the Prover}
\label{sec:preprocessor_prover}

@@

\subsection{The Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
\subsubsection{Handling output inside the loop body} Recall that the reducer program can perform the ouput command $out(e)$ inside the loop body $s_2$. In this case, we remember the output value in a fresh variable $v$ and replace the $out(e)$ command with an assignment $v:=e$.
Notice that the value of $v$ stays the same from the point it has been assigned a value until the end of the loop body $s_2$.
If the values of $v$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output values produced inside the loop body $s_2$.

@@

\hide{	
	\begin{figure}[t]
		\begin{minipage}{0.4\textwidth}
			\begin{algorithm}[H]
				\;\;$m := \cur + 10$; \;
				\Loop{}{
					$t:=\cur$\;
					\uIf{ $t> m$}{
						$m := t$ \;
					}
				};
				$out(m)$\;\;
			\end{algorithm}
			\caption*{(a) Reducer max$^+$}
		\end{minipage}
		\begin{minipage}{0.6\textwidth}
			\begin{algorithm}[H]
				$s:=*;$\;
				\Loop{}{
					\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
					\uElse{
						$t:=\cur$\;
						\uIf{ $t > m$}{
							$m := t$ \;
						}
					}
				};
				$out(m)$
			\end{algorithm}
			\caption*{(b) Reducer max$^{+\mathtt{fix}}$}
		\end{minipage}
		\caption{Reducer examples}
		\label{fig:reducer_max}
	\end{figure}
}

@@

\hide{
	\begin{wrapfigure}{r}{9cm}
		\centering
		\begin{minipage}{0.8\textwidth}

			\begin{algorithm}[H]
				$s'_0;s:=*;$\;
				\Loop{}{
					\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
					\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
					\lElseIf{ $s = 3$}{$\ldots$}
					\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
					\lElse{ $s_2$}
				};$s_3$\;
			\end{algorithm}
		\end{minipage}
		\caption{The second transformation task}
		\label{fig:general_program_transformation}
\end{wrapfigure}}

@@

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
After the transformation, the reducer $s_1;\rloop\{s_2\};s_3$ never calls the $\cur$ function before enthering the loop (the $s_1$ part) and never output values inside the loop body (the $s_2$ part). The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

\hide{

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;\textbf{3};\textbf{4};5] and any other input with two consecutive values being swapped, e.g., [1;2;4;3;5].
The swap operation can be applied arbitrarily many times to produce all possible permutations of the input.
It follows that that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;3;4;5], [1;2;4;3;5], [1;4;2;3;5], [4;1;2;3;5], etc.
Recall that the first program transformation task stores all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to prove the reducer $s_1;\rloop\{s_2\};s_3$ commutative. Notice that $s_1$ does not read from the input (assumed at the beginning of this subsection). The execution of $s_1$ is deterministic in the sense that from the same initial valuation, it outputs the same sequence of values and ends at the same valuation. For the execution of $\rloop\{s_2\}$, from the validity of Equation~(\ref{eq:commu}), any permutation of input values leads the reducer to the same final valuation and the same sequence of outputs at the end of the loop.
Before executing the commands $s_3$, all inputs have been read (the loop exit condition) already. Similar to the case of $s_1$, $s_3$ does not read from the input and hence the execution of $s_3$ is deterministic. It follows that for all permutations of inputs, the reducer outputs the same sequence of values.}

\subsection{Optimization: Live Variable Analysis}
\label{section:optimizations}
Here we explain how live variable analysis is used in J-ReCoVer to improve the precision of commutativity checking.

@@

\hide{
\begin{figure}[hbt]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			$s:= 0$; \;
			\Loop{}{
				$t:=\cur$;
				$s:= s + t$\;
			};
			$out(s)$\;\;\;
		\end{algorithm}
		\caption{A commutative reducer with an invalid Equation~(\ref{eq:commu}).}
		\label{fig:reducer_opt}
	\end{minipage}
	\ \ \ \ \ \ \ \
	\begin{minipage}{0.5\textwidth}
		\LinesNumbered
		\centering
		\begin{minipage}{0.75\textwidth}
		\begin{algorithm}[H]
			$c_1;c_2;\ldots;c_i$\;
			\lIf{$g_1$}{$\ldots$}
			\lElseIf{ $g_2$}{$\ldots$}
			\lElseIf{ $g_3$}{$\ldots$}
			\lElseIf{ $g_m$}{$c_{i+1};\ldots ;c_n; out(v);\ldots$}
			\lElse{ $\ldots$}
		\end{algorithm}
	\end{minipage}
		\caption{The $s_3$ command with an $out(v)$ command in nested branches.}
		\label{fig:nested_out}
	\end{minipage}
\end{figure}

To handle such a situation, we need an algorithm to collect a set of variables whose value may affect the output.
Recall that the first program transformation task uses a set of variables $V_2=\{v_{out_k} \mid k \in [1,l]\}$, where $l$ is the number of output commands occurred in $s_2$, to collect output values occurred in the loop body $s_2$.
The set $V_2$ is exactly the set of variables that will affect the output inside the loop.}

@@

\hide{
Without loss of generality, we assume there is only one output command after the loop. Then the code after the loop (i.e., the $s_3$ part of the reducer) can only be in the form of Figure~\ref{fig:nested_out}.
We unfold all concatenation of commands, so now the commands $c_1\ldots c_n$ are either assignment commands $w:=e$ or branch command $\ite{g}{c'_1}{c'_2}$, and the only output command stays in the $m$-th layer of the branch command after $c_i$. For the case that $m=0$, the entire branch command in Lines 2-6 of Figure~\ref{fig:nested_out} is replaced with $out(v)$.

The data flow analysis algorithm works as follows.
A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember the variables that might flow to the output command $out(v)$ in $s_3$ (in Figure~\ref{fig:nested_out}).
The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$.
For the case that a command $c_j$ for $j\in [1,n]$ is being processed, we need to handle two different cases:
\begin{itemize}
	\item $c_j$ is an assignment $w:=e$ and $w\in V_3$: all variables in $e$ are added to $V_3$.
	\item $c_j$ is a branch $\ite{g}{c'_1}{c'_2}$: the algorithm starts from the same set $V_3$ and processes $c'_1$ and $c'_2$ individually and independently (in a similar manner) to obtain $V_3^1$ and $V_3^2$. After that, the algorithm takes their union $V_3^1\cup V_3^2$ as the new value of $V_3$.
\end{itemize}
After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs that happen in $s_3$.
Then we modify the Equation~(\ref{eq:commu}) to the one below
\begin{equation}
F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu2}
\end{equation}
The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
The use of Equation~(\ref{eq:commu2}) is critical to the success of our approach. }
\hide{
\subsection{Computing the $F$ function using symbolic execution with an efficient data structure}
\label{section:opt2}

The computation of $F$ in Equation~(\ref{eq:commu2}) can be done in several different ways. For example, one can construct a formula in the style of bounded model checking. Here we apply \emph{symbolic execution} to generate $F$. Below we first review the standard symbolic execution algorithm and then explain how we use it to construct the formula $F$.

First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer.
One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $v:= e$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
\begin{enumerate}
	\item If $c$ is $v:= e$, the next symbolic state is $\langle pc, val[v \rightarrow val(e) ], cmds \rangle$.
	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$.
\end{enumerate}

Here we lift the $val$ function from variables to expressions in the standard way and $val[v \rightarrow val(e) ]$ is obtained by reassigning the value of $v$ in $val$ to $val(e)$. The execution from a symbolic state terminates when its path condition is conflict or the commands to be executed become empty, i.e., all commands have been processed. So far it is standard.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% More concretely, for the verification of a MapReduce program, we propose to do
% it in two phases. First, ensuring if all reducers are commutative. If some
% reducers are non-commutative, modifying them to commutative ones. Usually, the
% modification is not a difficult task~\cite{xiao14mr}, and can be done without
% affecting their functionality and performance. For example, assume that the
% task of a reducer is to find the name of the person with highest score. Such a
% reducer is non-commutative when the input include two people with the same
% highest score. This reducer can be made commutative, by also comparing the ID
% number of people with the same score. Once all reducers are made commutative,
% we reduce the verification problem to sequential verification by fixing a
% scheduler. In this two-phase approach, the key enabling technique is an
% efficient procedure for checking reducer commutativity.

@@

% \begin{equation*}
% \begin{array}{rclcl}
% \multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and 
%   Constant}\\
% e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid
%   e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
% g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & 
%   \textmd{Guard}\\
% s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && 
%   \textmd{Command}\\
% p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
% \end{array}
% \end{equation*}

\begin{wrapfigure}{r}{3cm}
		\vspace{-0.8cm}

%		\begin{minipage}{0.32\textwidth}
%			\begin{algorithm}[H]
%				$m := \cur$; \;
%				\Loop{}{
%					$t:=\cur$;\;
%					\uIf{ $t > m$}{
%						$m := t$
%					}
%				};
%				$out(m)$\;
%			\end{algorithm}
%			\caption*{(a) max}
%		\end{minipage}
%		\hspace{-0.5cm}

	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
	\end{minipage}

% 		\begin{minipage}{0.4\textwidth}
% 			\begin{algorithm}[H]
% 				$a:=*;s := 0; c:=0$;\;
% 				\Loop{}{
% 					$s := s+(\cur-a)^2$; \;
% 					$c := c+1$
% 				};
% 				$o := s/c$;\;
% 				$out(o)$
% 			\end{algorithm}
% 			\caption*{(c) standard deviation}
% 		\end{minipage}

	\caption{A reducer that computes the average value.}
	\vspace{-0.8cm}
	\label{fig:reducer_example}
\end{wrapfigure}

For presentation purposes, we provide a sketch of a highly simplified language
for describing reducers. The language allows us to give examples of reducers and
to discuss our approach for checking their commutativity. Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, their combinations over basic
arithmetic operations, or a non-deterministically chosen integer value $*$. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until
the entire input list is consumed. An example of a reducer is shown in
Figure~\ref{fig:reducer_example}.

Some reducers use two (or more) top-level loops to compute the output, possibly
interleaved with some non-looping code. {\color{red} These loops are executed
sequentially, repeatedly iterating over the input list from its beginning.} For
example, for calculating the standard deviation, one first computes the average
of inputs and then uses it to compute the final result (using two passes over
the input list). In that case, we suggest to verify the reducer by first
partitioning it into two (or more) reducers, each containing a single top-level
loop, and then verifying these reducers separately.\footnote{\color{red}Nested
loops can be transformed into branches of the top-level loop.} The top-level
loops communicate through shared variables.  After the transformation, reducers
corresponding to the second top-level loop (and possibly further such loops)
will work with random values of the shared variable, which over-approximates the
original behaviour.

For instance, for the case of the standard deviation, the first reducer computes
the average, and the other reducer performs the rest of the computation. The
second reducer replaces the average with a non-deterministic value and thus
``over-approximates'' all possible outputs of the standard deviation
computation. Equivalently, it computes an output for all possible ``average''
values. However, this reducer is still commutative, and its commutativity can be
verified.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview of the J-ReCoVer tool} \label{sec:overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Hadoop, Java;

The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework.
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either through our web-interface or binary application installed on their own machine.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Preprocessor and the Prover} \label{sec:preprocessor_prover}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%=============================================================================
\subsection{The Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
%=============================================================================

\subsubsection{Handling output inside the loop body} Recall that the reducer program can perform the ouput command $out(e)$ inside the loop body $s_2$. In this case, we remember the output value in a fresh variable $v$ and replace the $out(e)$ command with an assignment $v:=e$.
Notice that the value of $v$ stays the same from the point it has been assigned a value until the end of the loop body $s_2$.
If the values of $v$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output values produced inside the loop body $s_2$.

@@

%	\begin{figure}[t]
%		\begin{minipage}{0.4\textwidth}
%			\begin{algorithm}[H]
%				\;\;$m := \cur + 10$; \;
%				\Loop{}{
%					$t:=\cur$\;
%					\uIf{ $t> m$}{
%						$m := t$ \;
%					}
%				};
%				$out(m)$\;\;
%			\end{algorithm}
%			\caption*{(a) Reducer max$^+$}
%		\end{minipage}
%		\begin{minipage}{0.6\textwidth}
%			\begin{algorithm}[H]
%				$s:=*;$\;
%				\Loop{}{
%					\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
%					\uElse{
%						$t:=\cur$\;
%						\uIf{ $t > m$}{
%							$m := t$ \;
%						}
%					}
%				};
%				$out(m)$
%			\end{algorithm}
%			\caption*{(b) Reducer max$^{+\mathtt{fix}}$}
%		\end{minipage}
%		\caption{Reducer examples}
%		\label{fig:reducer_max}
%	\end{figure}

@@

%	\begin{wrapfigure}{r}{9cm}
%		\centering
%		\begin{minipage}{0.8\textwidth}
%			
%			\begin{algorithm}[H]
%				$s'_0;s:=*;$\;
%				\Loop{}{
%					\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
%					\lElseIf{ $s =2$}{ $x_{i_2} := \cur; 
%                                           s'_2; s:=3$}
%					\lElseIf{ $s = 3$}{$\ldots$}
%					\lElseIf{ $s = m$}{$x_{i_m} := \cur; 
%                                          s'_m; s:=m+1$}
%					\lElse{ $s_2$}
%				};$s_3$\;
%			\end{algorithm}
%		\end{minipage}
%		\caption{The second transformation task}
%		\label{fig:general_program_transformation}
%        \end{wrapfigure}

@@

%=============================================================================
\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
%=============================================================================

After the transformation, the reducer $s_1;\rloop\{s_2\};s_3$ never calls the $\cur$ function before enthering the loop (the $s_1$ part) and never output values inside the loop body (the $s_2$ part). The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

% In the example of ``average'', the values of $s$ and $c$ are the same no
% matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in
% the swapped order.  So it computes the same value for the input
% [1;2;\textbf{3};\textbf{4};5] and any other input with two consecutive values
% being swapped, e.g., [1;2;4;3;5].  The swap operation can be applied
% arbitrarily many times to produce all possible permutations of the input.  It
% follows that that starting from the same variable valuation, the loop ends
% with the same valuation for any permutation of input values, e.g., the reducer
% will output the same value for [1;2;3;4;5], [1;2;4;3;5], [1;4;2;3;5],
% [4;1;2;3;5], etc.  Recall that the first program transformation task stores
% all output values in variables that will stay unchanged until the end of the
% loop. The validity of Equation~(\ref{eq:commu}) also implies that the
% execution of the loop produces the same sequence of outputs.
% 
% We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to
% prove the reducer $s_1;\rloop\{s_2\};s_3$ commutative. Notice that $s_1$ does
% not read from the input (assumed at the beginning of this subsection). The
% execution of $s_1$ is deterministic in the sense that from the same initial
% valuation, it outputs the same sequence of values and ends at the same
% valuation. For the execution of $\rloop\{s_2\}$, from the validity of
% Equation~(\ref{eq:commu}), any permutation of input values leads the reducer
% to the same final valuation and the same sequence of outputs at the end of the
% loop.  Before executing the commands $s_3$, all inputs have been read (the
% loop exit condition) already. Similar to the case of $s_1$, $s_3$ does not
% read from the input and hence the execution of $s_3$ is deterministic. It
% follows that for all permutations of inputs, the reducer outputs the same
% sequence of values.
%

%=============================================================================
\subsection{Optimization: Live Variable Analysis} \label{section:optimizations}
%=============================================================================

Here we explain how live variable analysis is used in J-ReCoVer to improve the precision of commutativity checking.

@@

%\begin{figure}[hbt]
%	\begin{minipage}{0.4\textwidth}
%		\begin{algorithm}[H]
%			$s:= 0$; \;
%			\Loop{}{
%				$t:=\cur$;
%				$s:= s + t$\;
%			};
%			$out(s)$\;\;\;
%		\end{algorithm}
%		\caption{A commutative reducer with an invalid 
%                   Equation~(\ref{eq:commu}).}
%		\label{fig:reducer_opt}
%	\end{minipage}
%	\ \ \ \ \ \ \ \
%	\begin{minipage}{0.5\textwidth}
%		\LinesNumbered
%		\centering
%		\begin{minipage}{0.75\textwidth}
%		\begin{algorithm}[H]
%			$c_1;c_2;\ldots;c_i$\;
%			\lIf{$g_1$}{$\ldots$}
%			\lElseIf{ $g_2$}{$\ldots$}
%			\lElseIf{ $g_3$}{$\ldots$}
%			\lElseIf{ $g_m$}{$c_{i+1};\ldots ;c_n; out(v);\ldots$}
%			\lElse{ $\ldots$}
%		\end{algorithm}
%	\end{minipage}
%		\caption{The $s_3$ command with an $out(v)$ command in 
%                   nested branches.}
%		\label{fig:nested_out}
%	\end{minipage}
%\end{figure}

% To handle such a situation, we need an algorithm to collect a set of variables
% whose value may affect the output.  Recall that the first program
% transformation task uses a set of variables $V_2=\{v_{out_k} \mid k \in
% [1,l]\}$, where $l$ is the number of output commands occurred in $s_2$, to
% collect output values occurred in the loop body $s_2$.  The set $V_2$ is
% exactly the set of variables that will affect the output inside the loop.

@@

% Without loss of generality, we assume there is only one output command after the loop. Then the code after the loop (i.e., the $s_3$ part of the reducer) can only be in the form of Figure~\ref{fig:nested_out}.
% We unfold all concatenation of commands, so now the commands $c_1\ldots c_n$ are either assignment commands $w:=e$ or branch command $\ite{g}{c'_1}{c'_2}$, and the only output command stays in the $m$-th layer of the branch command after $c_i$. For the case that $m=0$, the entire branch command in Lines 2-6 of Figure~\ref{fig:nested_out} is replaced with $out(v)$.
% 
% The data flow analysis algorithm works as follows.
% A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember the variables that might flow to the output command $out(v)$ in $s_3$ (in Figure~\ref{fig:nested_out}).
% The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$.
% For the case that a command $c_j$ for $j\in [1,n]$ is being processed, we need to handle two different cases:
% \begin{itemize}
% 	\item $c_j$ is an assignment $w:=e$ and $w\in V_3$: all variables in $e$ are added to $V_3$.
% 	\item $c_j$ is a branch $\ite{g}{c'_1}{c'_2}$: the algorithm starts from the same set $V_3$ and processes $c'_1$ and $c'_2$ individually and independently (in a similar manner) to obtain $V_3^1$ and $V_3^2$. After that, the algorithm takes their union $V_3^1\cup V_3^2$ as the new value of $V_3$.
% \end{itemize}
% After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs that happen in $s_3$.
% Then we modify the Equation~(\ref{eq:commu}) to the one below
% \begin{equation}
% F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
% \label{eq:commu2}
% \end{equation}
% The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
% The use of Equation~(\ref{eq:commu2}) is critical to the success of our approach.

% \subsection{Computing the $F$ function using symbolic execution with an efficient data structure}
% \label{section:opt2}
% 
% The computation of $F$ in Equation~(\ref{eq:commu2}) can be done in several different ways. For example, one can construct a formula in the style of bounded model checking. Here we apply \emph{symbolic execution} to generate $F$. Below we first review the standard symbolic execution algorithm and then explain how we use it to construct the formula $F$.
% 
% 
% First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
% A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer.
% One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $v:= e$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
% \begin{enumerate}
% 	\item If $c$ is $v:= e$, the next symbolic state is $\langle pc, val[v \rightarrow val(e) ], cmds \rangle$.
% 	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$.
% \end{enumerate}
% 
% Here we lift the $val$ function from variables to expressions in the standard way and $val[v \rightarrow val(e) ]$ is obtained by reassigning the value of $v$ in $val$ to $val(e)$. The execution from a symbolic state terminates when its path condition is conflict or the commands to be executed become empty, i.e., all commands have been processed. So far it is standard.
% 
% For the loop body $s_2$, symbolic execution is guaranteed to terminate because there exist no commands in $s_2$ that may lead to infinite computations.
% When the execution terminates, we collect all states with empty commands $\langle pc, val, \emptyset \rangle$ as a set $T$ and use them to create the function $F$ in Equation~(\ref{eq:commu2}).
% First, we create for each variable $x\in\Var$ a formula $f_x$ whose value is $\bigvee_{\langle pc, val, \emptyset \rangle \in T} (pc \wedge val(x))$.
% The function $F(n,x_1,x_2,\ldots,x_k)$ simply returns the tuple $(f_{x_1},f_{x_2},\ldots, f_{x_k})$.
% 
% However, doing symbolic execution in a naive way can be extremely inefficient.
